---
title: ""
author: ""
format:
  html:
    toc: true
    toc-expand: true
    toc-depth: 4
    toc-location: left
    number-sections: true
    self-contained: true
    code-fold: true
    output-file: "ESM_1"
editor_options: 
  chunk_output_type: console
execute: 
  warning: false
  message: false
  echo: true
---

# Project Setup

```{r}
#| label:  setup
library(emmeans)
library(corrplot)
library(multcomp)
library(FSA)
library(agricolae)
library(tidyverse)
library(car)
library(lme4)
library(googlesheets4)
library(factoextra)
library(FactoMineR)
source('https://inkaverse.com/setup.r')
```

# Data import

```{r}
gs <- "https://docs.google.com/spreadsheets/d/1GgfIOmTGyvjtuVQzJgFA8ueD1yqRh7x7gyueLdQVd90/edit?usp=sharing" %>% 
  as_sheets_id()

inoculation <- gs %>% 
  range_read(ss = .,sheet = "STable3") %>% 
   mutate(across(1:6, ~ as.factor(.)))

glimpse(inoculation)

climate <- gs %>% 
  range_read(ss = .,sheet = "STable4") %>% 
   mutate(across(1:2, ~ as.factor(.)))

glimpse(climate)
```

# Data summary

Summary of the number of data points recorded for each treatment and evaluated variable.

```{r}
sm <- inoculation %>% 
  group_by(Site, treat) %>% 
  summarise(across(pl_height:yield, ~ sum(!is.na(.))))

sm
```

# References

Broman, K. W., & Woo, K. H. (2017). Data organization in spreadsheets. The American Statistician, 72(1), 2–10. https://doi.org/10.1080/00031305.2017.1375989

Zuur, A. F., Ieno, E. N., & Elphick, C. S. (2009). A protocol for data exploration to avoid common statistical problems. Methods in Ecology and Evolution, 1(1), 3–14. https://doi.org/10.1111/j.2041-210x.2009.00001.x

Francoishusson. (2017, July 13). PCA course using FactoMineR | R-bloggers. R-bloggers. https://www.r-bloggers.com/2017/07/pca-course-using-factominer/

Kozak, M., & Piepho, H. (2017). What’s normal anyway? Residual plots are more telling than significance tests when checking ANOVA assumptions. Journal of Agronomy and Crop Science, 204(1), 86–98. https://doi.org/10.1111/jac.12220

Tanaka, E., & Hui, F. K. C. (2019). Symbolic formulae for linear mixed models. In Communications in computer and information science (pp. 3–21). https://doi.org/10.1007/978-981-15-1960-4_1

Schielzeth, H., Dingemanse, N. J., Nakagawa, S., Westneat, D. F., Allegue, H., Teplitsky, C., Réale, D., Dochtermann, N. A., Garamszegi, L. Z., & Araya‐Ajoy, Y. G. (2020). Robustness of linear mixed‐effects models to violations of distributional assumptions. Methods in Ecology and Evolution, 11(9), 1141–1152. https://doi.org/10.1111/2041-210x.13434

# Research Objective

1. To evaluate the effects of bio-inoculation and PEG 6000 stress on pea seed biomass, weight, and germination on pea crop.

> Figura: Proceso germinativo de arverja inoculados co bioestimulantes en condicioens de estres hídrico simulado. (a) Gerinación (%). (b) Tiempo medio de Germinación (c) Sincronia de la Germinación. (d) Indice de vigor de la germinación
(e) Peso se las radiculas

> PCA germinación

1. To demonstrate the combined effects of bio-inoculation and fertilization on field experiment on pea crop

- Correlación de las variables por ambiente
- Una MLT y calcular las medias ajustadas (BLUEs)

> Figura: Las variables de campo

> Figura: PCA

## RO 1

```{r}

```

## RO 2

```{r}

```
