---
title: ""
author: ""
format:
  html:
    toc: true
    toc-expand: true
    toc-depth: 4
    toc-location: left
    number-sections: true
    self-contained: true
    code-fold: true
    output-file: "ESM_1"
editor_options: 
  chunk_output_type: console
execute: 
  warning: false
  message: false
  echo: true
---

# Project Setup

```{r}
#| label:  setup
library(emmeans)
library(corrplot)
library(multcomp)
library(FSA)
library(agricolae)
library(tidyverse)
library(car)
library(lme4)
library(googlesheets4)
library(factoextra)
library(FactoMineR)
library(GerminaR)
source("https://inkaverse.com/setup.r")
```

# Data import

```{r}
gs <- "https://docs.google.com/spreadsheets/d/1GgfIOmTGyvjtuVQzJgFA8ueD1yqRh7x7gyueLdQVd90/edit?usp=sharing" %>%
  as_sheets_id()

germination <- gs %>%
  range_read(ss = ., sheet = "STable2") %>%
  mutate(across(1:9, ~ as.factor(.)))

glimpse(germination)

inoculation <- gs %>%
  range_read(ss = ., sheet = "STable3") %>%
  mutate(across(1:6, ~ as.factor(.)))

glimpse(inoculation)

climate <- gs %>%
  range_read(ss = ., sheet = "STable4") %>%
  mutate(across(1:2, ~ as.factor(.)))

glimpse(climate)
```

# Data summary

Summary of the number of data points recorded for each treatment and evaluated variable.

```{r}
sm <- germination %>%
  group_by(Inoculant) %>%
  summarise(across(starts_with("ger"), ~ sum(!is.na(.))))

sm %>% kable(align = "c")

sm <- inoculation %>%
  group_by(Site, treat) %>%
  summarise(across(pl_height:yield, ~ sum(!is.na(.))))

sm %>% kable(align = "c")
```

# Research Objective

1. To evaluate the effects of bio-inoculation and PEG 6000 stress on pea seed biomass, weight, and germination on pea crop.

> Figura: Proceso germinativo de arverja inoculados co bioestimulantes en condicioens de estres hídrico simulado. (a) Gerinación (%). (b) Tiempo medio de Germinación (c) Sincronia de la Germinación. (d) Indice de vigor de la germinación
(e) Peso se las radiculas

> PCA germinación

1. To demonstrate the combined effects of bio-inoculation and fertilization on field experiment on pea crop

- Correlación de las variables por ambiente
- Una MLT y calcular las medias ajustadas (BLUEs)

> Figura: Las variables de campo

> Figura: PCA

## RO 1

To evaluate the effects of bio-inoculation and PEG 6000 stress on pea seed biomass, weight, and germination on pea crop.

### Germination (%)

#### Standard germination

```{r}
fb <- germination %>%
  select(sort, Inoculant,peg, block, seeds, starts_with("ger"))

gerdr <- fb

prefix <- "ger_"
dtnames <- gerdr %>% names()
evalprf <- dtnames[grepl(prefix, dtnames)]

gercount <- gerdr %>% 
  select(starts_with(prefix)) %>% 
  rowwise() %>% 
  summarize(.data[[evalprf[1]]], d = list(diff(c_across(everything())))) %>% 
  unnest_wider(d, names_sep="", names_repair=\(x) evalprf) %>% 
  cbind(gerdr %>% select(!starts_with(prefix)), .) %>% 
  mutate(across(where(is.numeric), ~ ifelse(. < 0, 0, .)))
  
# gercount %>% web_table()
```

#### Performe ten germination indices

```{r}
gercount %>%
  head(10) %>%
  kable(caption = "Germination dataset", align = "c")

glimpse(gercount)

# germination analysis (ten variables)

gsm <- gercount %>%
  as.data.frame() %>% 
  ger_summary(SeedN = "seeds"
                   , evalName = "ger_"
                   , data = .
                   )

# Pea data set processed

gsm %>% 
  head(10) %>% 
  mutate(across(where(is.numeric), ~round(., 2))) %>% 
  kable(caption = "Function ger_summary performe ten germination indices")
```

#### Germination Percentage (GRP)

```{r}
# analysis of variance

av <- lmer(grp ~ 1 + (1|block) + Inoculant*peg, data = gsm)

Anova(av, type = 3, test.statistic = "F")

mc1 <- emmeans(av, ~ peg|Inoculant) %>%
  cld(Letters = letters, reversed = T) %>% 
  mutate(across(".group", ~ trimws(.))) %>% 
  rename(sig1 = ".group")

mc2 <- emmeans(av, ~ Inoculant|peg) %>%
  cld(Letters = letters, reversed = T) %>% 
  mutate(across(".group", ~ trimws(.))) %>% 
  mutate(across(".group", ~ toupper(.))) %>% 
  rename(sig2 = ".group")

mc <- merge(mc2, mc1) %>% 
  unite(col = "group", c("sig1", "sig2"), sep = "")

mc %>% kable(caption = "Germination percentage mean comparision")

grp <- mc %>% 
   fplot(data = .
       , type = "bar"
       , x = "Inoculant"
       , y = "emmean"
       , group = "peg"
       , ylimits = c(0, 120, 30)
       , ylab = "Germination ('%')"
       , xlab = "Inoculant"
       , glab = "PEG (MPa)"
       , error = "SE"
       , sig = "group"
       , color = F
       )

grp
```

#### Mean Germination Time (MGT)

```{r}
# analysis of variance

av <- lmer(mgt ~ 1 + (1|block) + Inoculant*peg, data = gsm)

Anova(av, type = 3, test.statistic = "F")

mc1 <- emmeans(av, ~ peg|Inoculant) %>%
  cld(Letters = letters, reversed = T) %>% 
  mutate(across(".group", ~ trimws(.))) %>% 
  rename(sig1 = ".group")

mc2 <- emmeans(av, ~ Inoculant|peg) %>%
  cld(Letters = letters, reversed = T) %>% 
  mutate(across(".group", ~ trimws(.))) %>% 
  mutate(across(".group", ~ toupper(.))) %>% 
  rename(sig2 = ".group")

mc <- merge(mc2, mc1) %>% 
  unite(col = "group", c("sig1", "sig2"), sep = "")

mc %>% kable(caption = "Germination percentage mean comparision")

grp <- mc %>% 
   fplot(data = .
       , type = "bar"
       , x = "Inoculant"
       , y = "emmean"
       , group = "peg"
       , ylimits = c(0, 3, 0.5)
       , ylab = "Mean germination time (days)"
       , xlab = "Inoculant"
       , glab = "PEG (MPa)"
       , error = "SE"
       , sig = "group"
       , color = F
       )

grp
```

#### Synchronization Index (SYN)

```{r}
# analysis of variance

av <- lmer(syn ~ 1 + (1|block) + Inoculant*peg, data = gsm)

Anova(av, type = 3, test.statistic = "F")

mc1 <- emmeans(av, ~ peg|Inoculant) %>%
  cld(Letters = letters, reversed = T) %>% 
  mutate(across(".group", ~ trimws(.))) %>% 
  rename(sig1 = ".group")

mc2 <- emmeans(av, ~ Inoculant|peg) %>%
  cld(Letters = letters, reversed = T) %>% 
  mutate(across(".group", ~ trimws(.))) %>% 
  mutate(across(".group", ~ toupper(.))) %>% 
  rename(sig2 = ".group")

mc <- merge(mc2, mc1) %>% 
  unite(col = "group", c("sig1", "sig2"), sep = "")

mc %>% kable(caption = "Germination percentage mean comparision")

grp <- mc %>% 
   fplot(data = .
       , type = "bar"
       , x = "Inoculant"
       , y = "emmean"
       , group = "peg"
       , ylimits = c(0, 1, 0.25)
       , ylab = "Synchronization Index"
       , xlab = "Inoculant"
       , glab = "PEG (MPa)"
       , error = "SE"
       , sig = "group"
       , color = F
       )

grp
```

#### Germination vigor index (GVI)

```{r}
# analysis of variance

weight_res <- germination %>%
  group_by(Inoculant, peg, block) %>%
  summarise(weight = mean(peso, na.rm = TRUE), .groups = "drop")

gsm2 <- gsm %>%
  left_join(weight_res, by = c("Inoculant", "peg", "block")) %>%
  mutate(gv = weight * grp/100) %>% 
   mutate(across(where(is.numeric), ~round(., 2)))

av <- lmer(gv ~ 1 + (1|block) + Inoculant*peg, data = gsm2)

Anova(av, type = 3, test.statistic = "F")

mc1 <- emmeans(av, ~ peg|Inoculant) %>%
  cld(Letters = letters, reversed = T) %>% 
  mutate(across(".group", ~ trimws(.))) %>% 
  rename(sig1 = ".group")

mc2 <- emmeans(av, ~ Inoculant|peg) %>%
  cld(Letters = letters, reversed = T) %>% 
  mutate(across(".group", ~ trimws(.))) %>% 
  mutate(across(".group", ~ toupper(.))) %>% 
  rename(sig2 = ".group")

mc <- merge(mc2, mc1) %>% 
  unite(col = "group", c("sig1", "sig2"), sep = "")

mc %>% kable(caption = "Germination percentage mean comparision")

grp <- mc %>% 
   fplot(data = .
       , type = "bar"
       , x = "Inoculant"
       , y = "emmean"
       , group = "peg"
       , ylimits = c(0, 6, 2)
       , ylab = "Germination vigor Index"
       , xlab = "Inoculant"
       , glab = "PEG (MPa)"
       , error = "SE"
       , sig = "group"
       , color = F
       )

grp
```



## RO 2

To demonstrate the combined effects of bio-inoculation and fertilization on field experiment on pea crop

```{r}
```



# References

Broman, K. W., & Woo, K. H. (2017). Data organization in spreadsheets. The American Statistician, 72(1), 2–10. https://doi.org/10.1080/00031305.2017.1375989

Zuur, A. F., Ieno, E. N., & Elphick, C. S. (2009). A protocol for data exploration to avoid common statistical problems. Methods in Ecology and Evolution, 1(1), 3–14. https://doi.org/10.1111/j.2041-210x.2009.00001.x

Francoishusson. (2017, July 13). PCA course using FactoMineR | R-bloggers. R-bloggers. https://www.r-bloggers.com/2017/07/pca-course-using-factominer/

Kozak, M., & Piepho, H. (2017). What’s normal anyway? Residual plots are more telling than significance tests when checking ANOVA assumptions. Journal of Agronomy and Crop Science, 204(1), 86–98. https://doi.org/10.1111/jac.12220

Tanaka, E., & Hui, F. K. C. (2019). Symbolic formulae for linear mixed models. In Communications in computer and information science (pp. 3–21). https://doi.org/10.1007/978-981-15-1960-4_1

Schielzeth, H., Dingemanse, N. J., Nakagawa, S., Westneat, D. F., Allegue, H., Teplitsky, C., Réale, D., Dochtermann, N. A., Garamszegi, L. Z., & Araya‐Ajoy, Y. G. (2020). Robustness of linear mixed‐effects models to violations of distributional assumptions. Methods in Ecology and Evolution, 11(9), 1141–1152. https://doi.org/10.1111/2041-210x.13434
